---
output:
  word_document: default
  html_document: default
---
```{R, message=FALSE, warning=FALSE}
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(caret)
library(rpart)
library(rattle)
library(RColorBrewer)
```

```{r}
parole<-read_csv("parole.csv")
```


```{r}
parole=parole%>% mutate(male=as_factor(as.character(male)))%>%
  mutate(male=fct_recode(male,
                           "male"="1",
                           "female"="0"))
```


```{r}
parole=parole%>% mutate(race=as_factor(as.character(race)))%>%
  mutate(race=fct_recode(race,
                           "white"="1",
                           "otherwise"="2"))
```

```{r}
parole=parole%>% mutate(state=as_factor(as.character(state)))%>%
  mutate(state=fct_recode(state,
                           "Other"="1",
                           "Kentucky"="2",
                           "Louisiana"="3",
                           "Virginia"="4"))
```

```{r}
parole=parole%>% mutate(crime=as_factor(as.character(crime)))%>%
  mutate(crime=fct_recode(crime,
                           "Other crime"="1",
                           "larceny"="2",
                           "drug-related"="3",
                           "driving-related"="4"))
```

```{r}
parole=parole%>% mutate(multiple.offenses=as_factor(as.character(multiple.offenses)))%>%
  mutate(multiple.offenses=fct_recode(multiple.offenses,
                           "multiple offenses"="1",
                           "otherwise"="0"))
```

```{r}
parole=parole%>% mutate(violator=as_factor(as.character(violator)))%>%
  mutate(violator=fct_recode(violator,
                           "violator"="1",
                           "not violator"="0"))
```

```{r}
set.seed(12345)
train.rows=createDataPartition(y=parole$violator,p=0.7,list=FALSE)
train=parole[train.rows,]
test=parole[-train.rows,]
```

```{r}
traintree=rpart(violator~.,train,method="class")
fancyRpartPlot(traintree)
```

40 year old parolee from Louisiana who served a 5 year prison sentence? 
If white: Not a violator - Because he is from Louisiana we go right, because he is white, we go left, which makes him a non-violaotor.
Other: Violator Because he is from Lousianan we go right, because he his not white we go right againbecause he served 5 years we go left, because he is over 30 it lead to Violator.

```{r}
printcp(traintree)
```

```{r}
traintree2 = prune(traintree,cp= traintree$cptable[which.min(traintree$cptable[,"xerror"]),"CP"])
```

The majority class is non-violators.

```{r}
traintreepred=predict(traintree,train,type="class")
head(traintreepred)
```

```{r}
confusionMatrix(traintreepred,train$violator,positive="violator")
```

```{r}
testpred=predict(traintree,test,type="class")
head(testpred)
```


```{r}
confusionMatrix(testpred,test$violator,positive="violator")
```

```{r}
Blood<-read_csv("Blood.csv")
```


```{r}
Blood=Blood%>% mutate(DonatedMarch=as_factor(as.character(DonatedMarch)))%>%
  mutate(DonatedMarch=fct_recode(DonatedMarch,
                           "Yes"="1",
                           "No"="0"))
```

```{r}
set.seed(1234)
train.rows=createDataPartition(y=Blood$DonatedMarch,p=0.7,list=FALSE)
train2=Blood[train.rows,]
test2=Blood[-train.rows,]
```

```{r}
traintreeBlood=rpart(DonatedMarch~.,train2,method="class")
fancyRpartPlot(traintreeBlood)
```


```{r}
printcp(traintreeBlood)
```


```{r}
traintreeBlood = prune(traintreeBlood,cp= traintreeBlood$cptable[which.min(traintreeBlood$cptable[,"xerror"]),"CP"])
```

```{r}
traintreeBloodpred=predict(traintreeBlood,train2,type="class")
head(traintreeBloodpred)
```

```{r}
confusionMatrix(traintreeBloodpred,train2$DonatedMarch,positive="Yes")
```

```{r}
testpredBlood=predict(traintreeBlood,test2,type="class")
head(testpredBlood)
```


```{r}
confusionMatrix(testpredBlood,test2$DonatedMarch,positive="Yes")
```

It looks like accuracy decreases from 81.3% to 75.45% between the training and testing data when I run these predictions. Sensitivity and specificity also go down. 