---
output:
  word_document: default
  html_document: default
---
```{R, message=FALSE, warning=FALSE}
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(cluster)
library(dendextend)
library(factoextra)
```

```{r}
trucks=read_csv("trucks.csv")
str(trucks)
summary(trucks)
```

```{r}
ggplot(trucks,aes(x=Distance,y=Speeding))+
  geom_point()
```

There does appear to be some natural clustering based on the ggplot visualization.

```{r}
trucks2=as.data.frame(trucks%>%select(-Driver_ID))
```
```{r}
trucks2=scale(trucks2)
summary(trucks2)
```

```{r}
set.seed(1234)
cluster1<-kmeans(trucks2,2)
```

```{r}
fviz_cluster(cluster1,trucks2)
```

Two clusters looks a little forced. It looks more like it should be four.

```{r}
set.seed(123)
fviz_nbclust(trucks2,kmeans,method="wss")
```

Look like possibly four is the best per this method.

```{r}
set.seed(123)
fviz_nbclust(trucks2,kmeans,method="silhouette")
```

This one ALSO looks like four is optimal! There is a consensus between the two methods.

```{r}
set.seed(1234)
cluster2<-kmeans(trucks2,4)
cluster2
fviz_cluster(cluster2,trucks2)
```

That was NOT how I expected this to look. Now it looks almost like maybe 5 would have been better? Cluster 1 (the salmon colored one) seems a little forced to me.

```{r}
wine=read_csv("wineprice.csv")
str(wine)
summary(wine)
```

```{r}
wine2=as.data.frame(wine%>%select(-Year, -FrancePop))
```
```{r}
wine2=scale(wine2)
summary(wine2)
```

```{r}
set.seed(123)
fviz_nbclust(wine2,kmeans,method="wss")
```

Somewhere around five or six is what it looks like here, I think.

```{r}
set.seed(123)
fviz_nbclust(wine2,kmeans,method="silhouette")
```

This method identifies five, so there is consensus.

```{r}
set.seed(1234)
cluster2<-kmeans(wine2,5)
cluster2
fviz_cluster(cluster2,wine2)
```

```{r}
m=c("average", "single", "complete", "ward")
names(m)=c("average", "single", "complete", "ward")

ac=function(x) {agnes(wine2, method=x)$ac}
  map_dbl(m,ac)
```

```{r}
hc=agnes(wine2,method="ward")
pltree(hc,cex=0.6,hang=-1,main="Agglomerative Dendogram")
```

```{r}
hc2=diana(wine2)
pltree(hc2,cex=0.6,hang=-1,main="Diversive Dendogram")
```

